---
title: "Walmart_Time_Series_Analysis"
output: pdf_document
date: "2022-11-15"
---

```{r}
# import dependencies

library(tidyr)
library(dplyr)
library(corrplot)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(lubridate)
library(forecast)
```


# Importing the dataset
```{r}

walmart_df <- read.csv("Dataset/walmart-sales-dataset-of-45stores.csv")
walmart_df

#walmart_df$Date <- as.Date(walmart_df$Date, format = "%d-%m-%Y")

# Seperating the Date
#walmart_df[,"Year"] <- format(walmart_df[,"Date"],"%Y")
#walmart_df[,"Month"] <- format(walmart_df[,"Date"],"%m")
#walmart_df[,"Day"] <- format(walmart_df[,"Date"],"%d")

# Dataset dimension
dim(walmart_df)

#The dataset contains 6435 rows and 11 columns

#Checking null vales
colSums(is.na(walmart_df))

#The dataset contains no missing values


summary(walmart_df)

```
# Visualization

```{r}

# Let's find out the distribution of the variables.

par(mfrow=c(3,2))
hist(walmart_df$Temperature, col = 'light blue', main = "Temperature")
hist(walmart_df$Fuel_Price, col = 'light blue', main = "Fuel Price")
hist(walmart_df$CPI, col = 'light blue', main = "CPI")
hist(walmart_df$Unemployment, col = 'light blue', main = "Unemployment")
hist(walmart_df$Store, col = 'light blue', main = "Store Size")
hist(walmart_df$Weekly_Sales, col = 'light blue', main = "Weekly_Sales")

```
```{r}
ggplot(data = walmart_df
       ,aes(x=Weekly_Sales)) + 
  geom_histogram(bins=50)
```

The distribution of sales is clearly significantly skewed to the right. In this situation, the log function is typically used to account for the distributional skew. In the case of conventional approaches like regression analysis, the distribution of noise included in the data is frequently believed to be a symmetric distribution such as a normal distribution, which is why the distribution bias needs to be corrected.


If our initial data does not fit the bell curve, we may log convert it to make it as "normal" as possible, enhancing the validity of the statistical analysis results. Our weekly sales data is clustered for the low weekly sales because it is skewed to the right. 


```{r}

hist(log(walmart_df$Weekly_Sales), col = 'gray',
     main = "weekly sales log transformed",
     xlab ='log(Weekly Sales)')

```
# Box plot distribution
```{r}
num_cols <- unlist(lapply(walmart_df, is.numeric))         # Identify numeric columns

data_numeric <- walmart_df[, num_cols]
data_numeric

boxplot(data_numeric$Store)
boxplot(data_numeric$Weekly_Sales)
boxplot(data_numeric$Holiday_Flag)
boxplot(data_numeric$Temperature)
boxplot(data_numeric$Fuel_Price)
boxplot(data_numeric$CPI)
boxplot(data_numeric$Unemployment)
```

# Correlation Matrix 

```{r}
corr = cor(walmart_df[, c(3:8)])
corr

corrplot(corr, method = 'number')
```
There is a moderate positive correlation between sales, fuel price and holidays, and negative correlation between sales, unemployment, CPI, and temperature. 

# Weekly sales distribution in differnt month

```{r}
ggplot(walmart_df, aes(x = Month,y = Weekly_Sales )) + 
  geom_col() +
  facet_wrap(~Year) + 
  ggtitle("Weekly Sales Distribution in Different Months")
```



```{r}
ggplot(walmart_df, aes(x=Store, y=Weekly_Sales, color=Holiday_Flag)) +
  geom_point() 
```




```{r}
ggplot(walmart_df, aes(x = Store,y = Weekly_Sales )) + 
  geom_col() +
  facet_wrap(~Year) + 
  ggtitle("Weekly Sales Distribution in Different Stores")
```

Which store has the maximum sales? 

```{r}
sales_store <- aggregate(Weekly_Sales ~ Store, data = walmart_df, sum)
sales_store <- arrange(sales_store, desc(Weekly_Sales))
sales_store
```

Store 20 has the maximum value/weekly sales among other stores. 

```{r}
sales_store$Store <- as.character(sales_store$Store)
sales_store$Store <- factor(sales_store$Store, levels=unique(sales_store$Store))

colnames(sales_store) <- c("Store","Weekly_Sales")
barplot(sales_store$Weekly_Sales,names=sales_store$Store,
        main="Weekly Sales by Store"
)

```

The highest sales are at store 20, with a total value of 301M. The revenue of shop 4 is 299M which considered as the second-largest store in terms of sales. Store 33 has the lowest sales which is around 37.17M. 

```{r}

#Creating Holidays Data dataframe
Holiday_date <- c("12-02-2010", "11-02-2011", "10-02-2012", "08-02-2013","10-09-2010", "09-09-2011", "07-09-2012", "06-09-2013","26-11-2010", "25-11-2011", "23-11-2012", "29- 11-2013","31-12-2010", "30-12-2011", "28-12-2012", "27-12-2013")

holidays <-c(rep("Super Bowl", 4), 
           rep("Labour Day", 4),
           rep("Thanksgiving", 4), 
           rep("Christmas", 4))
holidays

Holidays_Data <- data.frame(holidays,Holiday_date)
Holidays_Data

#merging both dataframes
walmart_df.2<-merge(walmart_df,Holidays_Data, by.x= "Date", by.y="Holiday_date", all.x = TRUE)
walmart_df.2

#Replacing null values in Event with No_Holiday 
walmart_df.2$holidays = as.character(walmart_df.2$holidays) 
walmart_df.2$holidays[is.na(walmart_df.2$holidays)]= "No_Holiday" 
head(walmart_df.2)

```

```{r}
Holiday_Sales<-aggregate(Weekly_Sales ~ Events, data = walmart_df.2, mean)
Holiday_Sales
```

We have the most sales in the Thanksgiving each year and super bowl each year. 




```{r}
ggplot(walmart_df, aes(Date, Weekly_Sales)) +
  geom_line() +
  labs(title = "Weekly Sales vs Dates") +
  theme_classic()
```
```{r}
# plotting timeseries
walmart.ts  = ts(walmart_df$Weekly_Sales, frequency = 52,
                 start = c(2010,2), end = c(2012,10))
plot(walmart.ts, xlab="Time/Year", 
     ylab="Weekly Sales", bty='l')

```
The time series start in Feburary 2010, and ends in October 2012, and has frequency of 52 weeks per year. 

```{r}

walmart.lm <- tslm(walmart.ts ~ trend + I(trend^2))
par(mfrow = c(1,1))
plot(walmart.ts, xlab="Time/Year", ylab="Weekly Sales", bty="l")
lines(walmart.lm$fitted, lwd=2)

```

